### 2019-07-29 submission/manuscript.tex
Responding to edits & comments from Pat

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether PPIs promote CDIs using a CDI mouse model.
Mice were treated with a high daily dose of the PPI omeprazole for 7
days prior to \emph{C. difficile} challenge and were compared to mice
that were treated with the antibiotic clindamycin or both omeprazole and
clindamycin. High-dose PPI treatment was maintained throughout the
experiment, which included pre-treatment, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that omeprazole was not
sufficient to promote \emph{C. difficile} colonization. When omeprazole
treatment was combined with the antibiotic,one cage of mice remained
resistant to \emph{C. difficile} colonization, while the other cage was
colonized. 16S rRNA sequencing analysis revealed that omeprazole had
minimal impact on the murine microbiota throughout the 16 days of PPI
exposure. These results suggest PPI treatment alone is not sufficient to
disrupt microbiota resistance to \emph{C. difficle} infection in mice
that are normally resistant in the absence of antibiotic treatment.

@@

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but other factors may also contribute to CDI
incidence and recurrence. Interestingly, other medications besides
antibiotics can impact the microbiota. Proton pump inhibitors (PPIs) are
associated with alterations in the human intestinal microbiota through
observational and interventional studies and PPI use has also been
associated with CDI incidence and recurrence in epidemiological studies.
We evaluated the ability of a high dose of the PPI omeprazole to alter
the murine intestinal microbiota and disrupt colonization resistance to
\emph{C. difficile}. We found daily PPI treatment had minimal impact on
the murine fecal microbiota and was not enough to promote \emph{C.
difficile} colonization after challenge. Further studies are needed to
determine whether other factors such as the composition of the starting
bacterial community, comorbidities, and use of additional medications
contribute to the association between PPIs and CDIs seen in humans.

@@

\subsection{Introduction}\label{introduction}

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies as well as interventional clinical trials that have demonstrated
specific alterations in the intestinal microbiome were associated with
PPI use (4, 6). PPI-associated microbiota changes have been attributed
to the ability of PPIs to increase stomach acid pH which may promote the
survival of oral and pathogenic bacteria (4, 6). Human fecal microbiota
changes with PPI use include increases in Enterococcaceae,
Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae
and decreases in Ruminococcaceae (6).\\
Unfortunately, most of the studies suggesting a link between PPIs and
\emph{C. difficile} were retrospective and did not evaluate the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use play a role in the
association between PPIs and CDI incidence or recurrence. Additionally,
epidemiological studies have a limited capacity to address potential
confounders and comorbidities in patients that were on PPIs and
developed CDIs or recurrent CDIs (2, 5). Here, we evaluated the impact
of a daily high dose PPI treatment on the murine microbiome and
susceptibility to \emph{C. difficile} colonization in relation to
clindamycin, an antibiotic that perturbs the microbiome enough to allow
\emph{C. difficile} to colonize but is mild enough that \emph{C.
difficile} is cleared within 10 days (7).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test if PPI treatment alters the microbiome and promotes
susceptibility to CDIs, we gavaged mice daily with a high dose of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
The fecal bacterial communities of PPI-treated mice were compared to a
group of mice that received the antibiotic, clindamycin, and a group
that received both PPI and clindamycin treatment using 16S rRNA gene
sequencing. A principle coordinates analysis (PCoA) of the Bray-Curtis
distances over the initial 7 days of treatment revealed the bacterial
communities of PPI-treated mice remained relatively unchanged (Figure
1B). Enterococcaceae, Lactobacillaceae, Micrococcaceae,
Staphylococcaceae, Streptococcaceae, and Ruminococcaceae are all familes
that have previously been impacted by PPIs in human studies, but we
observed no significant fluctuations in the stool of our PPI-treated
mice throughout the course of the 16-day experiment (Figure 1C, Figure
S1). In contrast, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that high dose PPI treatment alone had no
significant impact on the murine fecal bacterial community.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether high dose PPI
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After 7 days of PPI treatment or 1 day after clindamycin
treatment, mice were challenged with \emph{C. difficile} 630 spores.
While all 4 of the clindamycin-treated mice were colonized with \emph{C.
difficile}, all of the PPI-treated mice were resistant to \emph{C.
difficile} colonization (Figure 2A). Interestingly, only 1 cage of the
PPI and clindamycin-treated mice were colonized, while the other cage
was resistant (Figure 2A). PCoAs of the bacterial communities from the 3
groups of mice after \emph{C. difficile} challenge, showed that the
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had
cleared \emph{C. difficile}, suggesting there was no difference in rate
of clearance for the mice that were initially colonized with \emph{C.
difficile} despite continuing to receive PPI treatment throughout the
course of the experiment. Our results suggest that PPI treatment alone
had no effects on bacterial community resistance to \emph{C. difficile}
colonization in mice. Instead most of the differences between our 3
treatment groups appear to be driven by clindamycin administration and
included decreased \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered
in clindamycin-treated mice that were challenged with \emph{C.
difficile} (1). These findings demonstrated that high dose PPI treatment
did not promote susceptibility to \emph{C. difficile} colonization.

\subsection{Discussion and
Conclusions}\label{discussion-and-conclusions}

Our findings that PPI treatment had minimal impact on the fecal
microbiome were comparable to another PPI mouse study that indicated
PPIs had more of an effect on the small intestinal microbiota compared
to the fecal microbiota (8). The same group demonstrated PPI treatment
increased the stomach pH in mice (8), which may improve survival of
bacteria passing through the stomach. We did not find significant
changes for the taxa observed to be significantly impacted by PPI use in
human studies. However, 3 of the families that typically increase were
absent or at low abundance in the mice from the beginning and may be one
contributing factor to our observation that PPIs had no significant
effects on bacterial taxa. Additionally, some of the significance of PPI
associations in human interventional trials appears to be driven by a
handful of specific taxa with overall differences on a PCoA plot
difficult to distinguish (9). Also, most of the studies to date compared
different individuals (PPI users versus non-users) or the same
individuals before and after treatment (4, 6), unfortunately these
limited microbiota snapshots may be skewed by day-to-day microbiota
variations (11) that would be revealed with more frequent longitudinal
sampling.\\
ALthough there have been a few \emph{C. difficile} mouse model studies

@@

additional antibiotic treatment(12--14), there were some key

@@

difficile} in mice not treated with antibiotics (12, 13). The presence
of \emph{C. difficile} in mice that were not treated with either
antibiotics or PPIs could be attributed to the higher dose of vegetative
cells (10\textsuperscript{8} CFU) used to challenge the mice and may
partly explain the observation of \emph{C. difficile} in PPI-treated
mice (12, 13). We have previously shown mice from our colony that were
not given antibiotics were resistant to \emph{C. difficile} 630 when
challenged with 10\textsuperscript{3} spores (15). The other group also
used a higher dose of 10\textsuperscript{6} spores or
10\textsuperscript{7} vegetative cells to challenge antibiotic-treated
mice or mice treated with both the PPI esomeprazole (40mg/kg dose) for 2
days and antibiotics and demonstrated the antibiotic/PPI-treated mice
developed more severe CDIs (14).\\
Our study extended previous work examining PPIs and \emph{C. difficile}
in mice to examine the contribution of the intestinal microbiota. We
found the PPI omeprazole had minimal impact on the murine intestinal
microbiota and in contrast to previous work, PPIs did not promote
\emph{C. difficile} colonization. 16S rRNA sequencing suggested that

@@

C57BL/6 mouse colony. These genera could be important contributors to
the associations between PPIs and CDIs in humans, and could be a
contributing factor to our observation that PPI treatment had no effect
on \emph{C. difficile} colonization in our CDI mouse model. There could
also be differences at the species level for taxa that are altered with
PPIs in humans compared to mice. Other murine starting communities or
the addition of human PPI treatment-associated strains may be needed to
determine whether PPIs can alter taxa and subsequently promote CDIs.\\
Factors such as age, body mass index, comorbidities, and use of other
medications in human studies may also be contributing to the association
between PPIs and CDI incidence or recurrence. For example, NSAID
treatment can increase CDI severity in mice (16) and PPIs are sometimes
used in combination with NSAIDs (3). This study addressed the impact of
PPIs with or without antibiotics on a murine model of CDI, and found
PPIs did not promote \emph{C. difficile} colonization. Future studies
are needed to determine whether age, other comorbidities and bacterial
strains that are less common in mice can increase the risk of CDIs or
recurrent CDIs when combined with PPI treatment.

@@

\textbf{Animals} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was spit between 2 cages with 2-3 mice housed per cage and male and
female mice were housed separately. All animal experiments were approved
by the University of Michigan Animal Care and Use Committee (IACUC)
under protocol number PRO00006983.

@@

\textbf{16S rRNA gene sequencing} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The V4 hypervariable region of the 16S rRNA gene was
amplified with Accuprime Pfx DNA polymerase (Thermo Fisher Scientific)
using previously described custom barcoded primers (17). The 16S rRNA
amplicon library was sequenced with the MiSeq (Illumina). Amplicons were
cleaned up and normalized with the SequalPrep Normalization Plate Kit
(ThermoFisher Scientific) and pooled amplicons were quantified with the
KAPA library quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis}Mothur (v1.40.5) was used for
all sequence processing steps (18) using the previously published
protocol (17). R (v.3.5.1) was used to generate figures and perform
statistical analysis.

\textbf{Statistical Analysis} To assess for differences in relative
abundance in families and genera across our 3 different treatment groups
(Clindamycin, Clindamycin + PPI, and PPI), we used a Kruskal-Wallis test
with a Benjamini-Hochberg correction for multiple comparisons.

@@

\textbf{Data availability} The 16S rRNA sequencing data have been
deposited in the NCBI Sequence Read Archive (Accession no.
\_\_\_\_\_\_\_\_\_)

@@

\hypertarget{ref-Jenior2018}{}
7. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
8. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
9. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
10. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-LlornsRico2019}{}
11. \textbf{Lloréns-Rico V}, \textbf{Raes J}. 2019. Tracking humans and
microbes. Nature \textbf{569}:632--633.
doi:\href{https://doi.org/10.1038/d41586-019-01591-y}{10.1038/d41586-019-01591-y}.

\hypertarget{ref-Kaur2007}{}
12. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
13. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
14. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
15. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Maseda2019}{}
16. \textbf{Maseda D}, \textbf{Zackular JP}, \textbf{Trindade B},
\textbf{Kirk L}, \textbf{Roxas JL}, \textbf{Rogers LM},
\textbf{Washington MK}, \textbf{Du L}, \textbf{Koyama T},
\textbf{Viswanathan VK}, \textbf{Vedantam G}, \textbf{Schloss PD},
\textbf{Crofford LJ}, \textbf{Skaar EP}, \textbf{Aronoff DM}. 2019.
Nonsteroidal anti-inflammatory drugs alter the microbiota and exacerbate
clostridium difficile colitis while dysregulating the inflammatory
response. mBio \textbf{10}.
doi:\href{https://doi.org/10.1128/mbio.02282-18}{10.1128/mbio.02282-18}.

\hypertarget{ref-Kozich2013}{}
17. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

>>>>

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether PPIs promote CDIs in a mouse model. Mice
receiving the PPI treatment were gavaged with 40 mg/kg of omeprazole
during a 7-day pretreatment phase, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that mice treated with
omeprazole were not colonized by \emph{C. difficile}. When omeprazole
treatment was combined with a single clindamycin treatment, one cage of
mice remained resistant to \emph{C. difficile} colonization, while the
other cage was colonized. Treatming mice with clindamycin and no
omeprazole followed by challenge resulted in \emph{C. difficile}
colonization. 16S rRNA gene sequencing analysis revealed that omeprazole
had minimal impact on the structure of the murine microbiota throughout
the 16 days of PPI exposure. These results suggest PPI treatment alone
is not sufficient to disrupt microbiota resistance to \emph{C. difficle}
infection in mice that are normally resistant in the absence of
antibiotic treatment.

@@

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but other factors may also contribute. In
epidemiological studies, PPI use has been associated with CDI incidence
and recurrence. Proton pump inhibitors (PPIs) have also been associated
with alterations in the human intestinal microbiota in observational and
interventional studies. We evaluated the effects of the PPI omeprazole
on the structure of the murine intestinal microbiota and its ability to
disrupt colonization resistance to \emph{C. difficile}. We found the PPI
treatment had minimal impact on the murine fecal microbiota and did not
promote \emph{C. difficile} colonization. Further studies are needed to
determine whether other factors such as the composition of the starting
bacterial community, comorbidities, and use of additional medications
contribute to the association between PPIs and CDIs seen in humans or
whether aspects of murine physiology may limit its utility to test these
types of hypotheses.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies as well as interventional clinical trials that have demonstrated
that specific alterations in the intestinal microbiome were associated
with PPI use (4, 6). PPI-associated microbiota changes have been
attributed to the ability of PPIs to increase stomach acid pH which may
promote the survival of oral and pathogenic bacteria (4, 6). Human fecal
microbiota changes with PPI use include increases in Enterococcaceae,
Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae
and decreases in Ruminococcaceae (6, 7). Several of these taxa have been
associated with C. difficile colonization in humans (8)\\
Unfortunately, most of the studies suggesting a link between PPIs and
\emph{C. difficile} were retrospective and did not evaluate the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of a daily high dose PPI treatment on the
murine microbiome and susceptibility to \emph{C. difficile} colonization
in relation to clindamycin, an antibiotic that perturbs the microbiome
enough to allow \emph{C. difficile} to colonize but is mild enough that
\emph{C. difficile} is cleared within 10 days (9).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test whether PPI treatment alters the microbiome and
promotes susceptibility to CDIs, we gavaged mice with 40 mg/kg of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
A principle coordinates analysis (PCoA) of the Bray-Curtis distances
over the initial 7 days of treatment revealed the bacterial communities
of PPI-treated mice remained relatively unchanged relative to animals
that did not recieve the PPI (Figure 1B). We observed no significant
fluctuations in the relative abundance of those taxa previously shown to
respond to PPI treatment throughout the course of the 16-day experiment
(Figure 1C, Figure S1A-B). We also observed no significant fluctuations
in relative abundances at the family and genus level over the beginning,
day of \emph{C. difficile} challenge, and end of the experiment for the
PPI-treated mice (Figure 1D, Figure S1C). Next, the fecal bacterial
communities of PPI-treated mice were compared to a group of mice that
received a single 10 mg/kg dose of clindamycin and a group that received
both PPI and clindamycin. In contrast to the dynamics of mice only
receiving omeprazole, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that the PPI treatment alone had a minimal
impact on the murine fecal bacterial community relative to the impact of
clindamycin treatment.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether PPI treatment
altered susceptibility to \emph{C. difficile} infection in mice. After
PPI treatment or clindamycin treatment, mice were challenged with
\emph{C. difficile} 630 spores. While \emph{C. difficile} colonized the
clindamycin treated mice, it was unable to colonize the PPI-treated mice
(Figure 2A). Interestingly, only 1 cage of mice that received both the
PPI and clindamycin were colonized, while the mice in other cage wwere
resistant (Figure 2A). The greatest shifts in bacterial communities
occurred in the clindamycin-treated mice (Figure 2B, Figure S2).
Regardless of whether the mice became colonized, all of the mice had
cleared \emph{C. difficile} within 5 days (Figure 2A), suggesting that
omeprazole did not affect the rate of clearance. Our results suggest
that PPI treatment had no effects on bacterial community resistance to
\emph{C. difficile} colonization in mice. Instead most of the
differences between our 3 treatment groups appear to be driven by
clindamycin administration (Figure 2C-D) and included previously
described decreases in \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae (1). These findings demonstrated that
high dose PPI treatment did not promote susceptibility to \emph{C.
difficile} colonization.

\subsection{Conclusions}\label{conclusions}

We found the PPI omeprazole did not impact the gut microbiota and did
not promote \emph{C. difficile} infection in mice. Our findings that PPI
treatment had minimal impact on the fecal microbiome were comparable to
another PPI mouse study that indicated PPIs had more of an effect on the
small intestinal microbiota compared to the fecal microbiota (10). The
same group demonstrated PPI treatment increased the stomach pH in mice
(10), which may improve survival of bacteria passing through the
stomach. We did not find significant changes for the taxa observed to be
significantly impacted by PPI use in human studies. However, 3 of the
human-associated taxa were absent or at low abundance in our mice.
Additionally, some of the significance of PPI associations in human
interventional trials appears to be driven by a handful of specific taxa
with overall differences on a PCoA plot difficult to distinguish (11).
The PPI-treated mice fecal communities did not change overall and no
differences in the relative abundances of individual families or genera
were observed over time. While our fecal microbiota findings are
comparable to what's been shown in another mouse study, whether
PPI-induced changes in specific bacterial abundances observed in humans
play a role in CDIs remains to be determined.\\
Although there have been a few \emph{C. difficile} mouse model studies

@@

additional antibiotic treatment (13--15), there were some key

@@

difficile} in mice not treated with antibiotics (13, 14). The presence
of \emph{C. difficile} in mice not treated with antibiotics may be
partly attributed to the higher dose of vegetative cells
(10\textsuperscript{8} CFU) used, but nevertheless makes it difficult to
interpret the extent to which PPI treatment contributed to \emph{C.
difficile} colonization in their mouse model (13, 14). In contrast, we
have previously shown mice from our colony that were not given
antibiotics were resistant to \emph{C. difficile} 630 when challenged
with 10\textsuperscript{3} spores (16) and previous work by our
collaborators demonstrated mice not given antibiotics were resistant to
10\textsuperscript{5} vegetative cells of \emph{C. difficile } VPI 10463
(17). The other PPI CDI mouse study also used a higher dose of
10\textsuperscript{6} spores or 10\textsuperscript{7} vegetative cells
to challenge antibiotic-treated mice or mice treated with both
antibiotics and the PPI esomeprazole (40mg/kg dose) for 2 days and
demonstrated the antibiotic/PPI-treated mice developed more severe CDIs
(15). We tested the same high 40 mg/kg PPI dose and expanded
pre-treatment to 7 days before challenge to test the impact of PPI
treatment alone on our CDI mouse model.\\
Our study extended the previous work examining PPIs and \emph{C.
difficile} in mice by evaluating the contribution of the intestinal
microbiota. We found the PPI omeprazole had no significant impact on
bacterial taxa within the murine intestinal microbiota over the 16-day
experiment and in contrast to previous work, PPIs did not alter \emph{C.
difficile} colonization resistance. 16S rRNA sequencing suggested that

@@

C57BL/6 mouse colony and variation in these 2 genera have been observed
across other facilities and vendors (18, 19). These genera could be
important contributors to the associations between PPIs and CDIs in
humans, and could be a contributing factor to our observation that PPI
treatment had no effect on \emph{C. difficile} colonization in our CDI
mouse model.\\
Beyond microbiome differences, factors such as age, body mass index,
comorbidities, and use of other medications in human studies may also be
contributing to the association between PPIs and CDI incidence or
recurrence. This study addressed the impact of PPIs with or without
antibiotics on a murine model of CDI, and found PPIs did not promote
\emph{C. difficile} colonization. The epidemiological evidence linking
PPIs to CDIs is primarily from observational studies, which makes
determining causality and whether other risk factors play a role
challenging (20). Future studies are needed to determine whether age,
other comorbidities and bacterial strains that are less common in mice
can increase the risk of CDIs or recurrent CDIs when combined with PPI
treatment.

@@

\textbf{Animals} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was spit between 2 cages with 2-3 mice housed per cage and male and
female mice housed separately. All animal experiments were approved by
the University of Michigan Animal Care and Use Committee (IACUC) under
protocol number PRO00006983.

@@

\textbf{16S rRNA gene sequencing} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The ZymoBIOMICS\textsuperscript{TM} Microbial
Community DNA Standard (Zymo, CA, USA) was used as a mock community (21)
and water was used as a negative control. The V4 hypervariable region of
the 16S rRNA gene was amplified with Accuprime Pfx DNA polymerase
(Thermo Fisher Scientific) using previously described custom barcoded
primers (22). The 16S rRNA amplicon library was sequenced with the MiSeq
(Illumina). Amplicons were cleaned up and normalized with the SequalPrep
Normalization Plate Kit (ThermoFisher Scientific) and pooled amplicons
were quantified with the KAPA library quantification kit (KAPA
Biosystems).

\textbf{16S rRNA gene sequence analysis}Mothur (v1.40.5) was used for
all sequence processing steps (23) using the previously published
protocol (22). In brief, forward and reverse reads for each sample were
combined and low-quality sequences and chimeras were removed. Duplicate
sequences were merged, before taxonomy assignment using a modified
version (v16) of the Ribosomal Database Project reference database
(v11.5) with an 80\% cutoff. Operational taxonomic units (OTUs) were
assigned with the opticlust clustering algorithm using a 97\% similarity
threshold. To adjust for uneven sequencing across samples, all samples
were rarefied to 3000 sequences, 1,000 times. PCoAs were generated based
on Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

\textbf{Statistical Analysis} To test for differences in relative
abundances in families and genera across our 3 different treatment
groups (Clindamycin, Clindamycin + PPI, and PPI) or within the PPI
treatment group across 3 timepoints (Day -7, 0, and 9), we used a
Kruskal-Wallis test with a Benjamini-Hochberg correction for multiple
comparisons.

@@

\textbf{Data availability} The 16S rRNA sequencing data have been
deposited in the NCBI Sequence Read Archive (Accession no. PRJNA554866).

@@

\hypertarget{ref-maier2018extensive}{}
7. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Schuberte01021-14}{}
8. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mBio.01021-14}{10.1128/mBio.01021-14}.

\hypertarget{ref-Jenior2018}{}
9. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
10. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
11. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
12. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-Kaur2007}{}
13. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
14. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
15. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
16. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Reeves2011}{}
17. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model ofClostridium difficileInfection. Gut Microbes
\textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.
